---
title: 'Thesis title'
shorttitle        : "shortitle"
author: 'Author name'
date: 'date'
year: 'year'
institution: 'University of Guelph'
advisor: 'advisor name'
#altadvisor: 'MJ' 
department: 'department'
degree: 'degree'
field: 'field (if necessary)' #comment out if unecessary 
#specialization: 'Statistics'
knit: bookdown::render_book
site: bookdown::bookdown_site
#include additional LaTeX code below if necessary 
header-includes:
#University of Guelph formatting is only supported pdf output 
output:
  thesisdown::thesis_pdf: 
    pandoc_args: ["--biblatex"]
    includes:
     after_body: ['80-references.Rmd', '81-appendix.Rmd'] #add any additional files to be compiled after the main text
# access the content in 00-abstract.Rmd file to print the abstract. 
abstract: '`r if(knitr:::is_latex_output()) paste(readLines(list.files(path = here::here(),pattern = "00--abstract.Rmd", recursive = TRUE, full.names = TRUE)), collapse = "\n  ")`'
# The acknowledgements section can also be accessed like the abstract, or simply printed from the YAML header as shown below. Note that a tab 
# is needed on the line after the `|`.
acknowledgements: |
  [Acknowledgements go here]
#comment out dedication if not necessary
dedication: |
  [Dedication goes here if necessary] 
#abbreviations:
#  ABC: American Broadcasting Company
#  CBS: Colombia Broadcasting System
#  CUS: Computer User Services
#  PBS: Public Broadcasting System
# Specify the location of the bibliography below
bibliography: bib/references.bib
#if biblatex is not desired and you want to format references using .csl file, uncomment line below 
#csl: csl/apa.csl 
lot: true  #list of tables 
lof: true  #list of figures
loa: true  #list of appendices
toc-depth: '5' #header level depth for table of contents
linenumbers: false #whether to put linenumbers in text (effective when making drafts of dissertation)
draft: false #prints 'DRAFT' watermark on pages (useful when making drafts)
ArialFont: false #Times New Roman set as default if false
fontsize: '12pt' #options: 10pt, 11pt, or 12pt (does not have to be wrapped in quotation marks)
linespacing: 2
backref: true #include backreferences
linkcolor: blue #color of all hyperlinks; set to black for print 
print: false #creates print version of thesis with blank pages in preamble
print_refs: false #prints references (useful when drafting); use in conjunction with knit_exit() to stop rendering at any point 
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence them (add # before each line). 

If you have other LaTeX packages you would like to include, list the packages after hyphens on new lines in the header-includes section.
-->

```{r package_loading_int, include=F}
#load packages
library(easypackages)
packages <- c('tidyverse','guelphddown', 'kableExtra', 'knitr') 
libraries(packages)
devtools::load_all()
#needed to compile figures produces by create_apa7_table
knitr::opts_chunk$set(message = F, results = "asis")
options(tinytex.compile.min_times = 3)
#write_bib(x = packages, file = 'bib/package_references.bib') #run code once for including packages in a .bib file 
```

```{r eval=F, echo=F}
#text below can be used to insert citations in the references section that do not appear in the text
```

\nocite{R-tidyverse, R-nonlinSims, R-nonlinSimsAnalysis, R-devtools, R-RColorBrewer, R-cowplot, R-data.table, R-egg, R-ggbrace, R-ggtext, R-kableExtra, R-knitr}




# Introduction 
## Quotes

\begin{quote}
    ``Neither the behavior of human beings nor the activities of organizations can be defined without reference to time, and temporal aspects are critical for understanding them" \parencite[][p. 136]{navarro2015}.
\end{quote}

## Footnotes 

Despite the considerable attention given towards investigating processes over time and the ubiquity of time in organizational theory, the prevalence of longitudinal research has historically remained low. One study examined the prevalence of longitudinal research from 1970--2006 across five organizational psychology journals and found that 4% of articles used longitudinal designs [@roe2014b]. Another survey of two applied psychology journals in 2005 found that approximately 10% (10 of 105 studies) of studies used longitudinal designs [@roe2008]. Similarly, two surveys of studies employing longitudinal designs with mediation analysis found that, across five journals, only about 10% (7 of 72 studies) did so in 2005 [@maxwell2007] and approximately 16% (15 of 92 studies) did so in 2006 [@mitchell2013].\footnote{Note that the definition of a longitudinal design in \textcite{maxwell2007} and \textcite{mitchell2013} required that measurements be taken over at least three time points so that measurements of the predictor, mediator, and outcome variables were separated over time.} Thus, the prevalence of longitudinal research has remained low.

Conducting longitudinal research presents researchers with several challenges. Many challenges are those from cross-sectional research only amplified [for a review, see @bergman1993].\footnote{It should be noted that conducting a longitudinal study does alleviate some issues encountered in conducting cross-sectional research. For example, taking measurements over multiple time points likely reduces common method variance \parencites{podsakoff2003}[for an example, see ][]{ostroff2002}.} For example, greater efforts have to be made to prevent missing data which can increase over time [@newman2008; @dillman2014]. Likewise, the adverse effects of well-documented biases such as demand characteristics [@orne1962] and social desirability [@nederhof1985] have to be countered at each time point. Outside of challenges shared with cross-sectional research, conducting longitudinal research also presents new challenges. Analyses of longitudinal data have to consider complications such as how to model error structures [@grimm2010a], check for measurement non-invariance over time [the extent to which a construct is measured with the same measurement model over time\; @mellenbergh1989], and how to center/process data to appropriately answer research questions [@enders2007; @wang2015]. 

## Figures

```{r montecarlo, echo=F}
create_apa7_figure(orientation = "landscape", samepage = T, figure_bottom_margin = "0cm", 
                   title = "Depiction of Monte Carlo Method", figure_tag = "MonteCarlo-comparison", image_ratio = "0.7", 
                   figure_path = "figures/Monte_Carlo_comparison", 
                   footnote = "Comparison of inferential approach with the Monte Carlo approach. The inferential approach begins with a collected sample and then estimates the population parameter using an appropriate statistical model. The difference between the estimated and population value can be conceptualized as error. Because the population value is generally unknown in the inferential approach, it cannot estimate how much error is introduced by any given methodological or statistical deficiency. To estimate how much error is introduced by any given methodological or statistical deficiency, the Monte Carlo method needs to be used, which constitutes four steps. The Monte Carlo method first defines a population by setting parameter values. Second, many samples are generated from the pre-defined population, with some methodological deficiency built in to each data set (in this case, each sample has a specific amount of missing data). Third, each generated sample is then analyzed and the population estimates of each statistical model are averaged and compared to the pre-determined parameter value. Fourth, the difference between the estimate average and the known population value defines the extent to which the missing data manipulation affected parameter estimation (the difference between the population and average estimated population value is the parameter bias).")
```


## Tables

```{r systematicReviewCount, echo=F}
table_1 <-  data.frame('Effect' = c('\\textbf{Main effects}', 'Number of measurements (NM)', 'Spacing of measurements (SM)', 'Time structuredness (TS)', 'Sample size (S)', 
                                    '\\textbf{Two-way interactions}', 'NM x SM', 'NM x TS', 'NM x S', 'SM x TS', 'SM x S', 'TS x S',
                                    '\\textbf{Three-way interactions}', 'NM x SM x TS', 'NM x SM x S', 'NM x TS x S', 'SM x TS x S'), 
               'Linear pattern' = c('', '11 studies', '1 study', '2 studies','11 studies', '', '1 study',  '1 study', '9 studies', '\\textbf{Cell 2}', 
                                    '\\textbf{Cell 4}', '1 study', '', '\\textbf{Cell 6}', '\\textbf{Cell 8}', 
                                    '1 study', ' \\textbf{Cell 11}'),
               'Nonlinear pattern' = c('', '6 studies', '1 study', '1 study', '7 studies', '', '1 study', ' \\textbf{Cell 1 (\\hyperref[Exp3]{Exp. 3})}', '5 studies', 
                                       '\\textbf{Cell 3}', '\\textbf{Cell 5 (\\hyperref[Exp2]{Exp. 2})}' , '2 studies', '', ' \\textbf{Cell 7}',
                                       '\\textbf{Cell 9 (\\hyperref[Exp2]{Exp. 2})}', ' \\textbf{Cell 10 (\\hyperref[Exp3]{Exp. 3})}', '\\textbf{Cell 12}'), check.names = F)
kbl(x = table_1, booktabs = TRUE, format = 'latex', longtable = TRUE, 
    linesep = c('\\cmidrule{1-3}',
        rep(' ', times = 3), '\\cmidrule{1-3}', '\\cmidrule{1-3}', 
        rep(' ', times = 5), '\\cmidrule{1-3}', '\\cmidrule{1-3}',
        rep(' ', times = 3)), 
    align = c('l', 'c', 'c'), 
    caption = 'Number of Simulation Studies That Have Investigated Longitudinal Issues with Linear and Nonlinear Change Patterns (\\textit{n} = 17)', 
    escape=F) %>%
  column_spec(1, width = '5.5cm') %>%
   column_spec(2, background = c(rep('white', times = 9), 
                                rep('#E4E2E2', times = 1), 
                                '#E4E2E2',
                                'white', 'white',
                                rep('#E4E2E2', times = 1), 
                                rep('#E4E2E2', times = 1),
                                'white',
                                '#E4E2E2'), 
              width = '8cm') %>%
  column_spec(3, background = c(rep('white', times = 7), 
                                '#C7C4C4',
                                'white', 
                                '#C7C4C4',
                                '#C7C4C4',
                                'white', 'white', 
                                 rep('#C7C4C4', times = 1), 
                                 rep('#C7C4C4', times = 2),
                                 '#C7C4C4'),  
              width = '8cm') %>% 
       kable_styling(latex_options= c('hold_position', 'repeat_header'), position = 'left') %>%
 footnote(general_title = '', general = '\\\\textit{Note. }Cells are only numbered for effects that have not been investigated. Cells shaded in light and dark grey, respectively indicate effects that have not been investigated with linear and nonlinear  patterns of change.', 
          footnote_as_chunk = F, escape = F, threeparttable = T) %>%
  landscape(margin = '2.54cm')
```


```{r systematicReview, echo=F, include=T}
table_1 <-  data.frame(
  'Effect' = c('\\textbf{Main effects}', 'Number of measurements (NM)', 'Spacing of measurements (SM)', 'Time structuredness (TS)', 'Sample size (S)', 
               '\\textbf{Two-way interactions}', 'NM x SM', 'NM x TS', 'NM x S', 'SM x TS', 'SM x S', 'TS x S',
               '\\textbf{Three-way interactions}', 'NM x SM x TS', 'NM x SM x S', 'NM x TS x S', 'SM x TS x S'), 
                       
'Linear pattern' = c('', 
                     '\\parencites[][\\textsuperscript{a}]{timmons2015}[][\\textsuperscript{b}$^{\\mho}$]{murphy2011}[][\\textsuperscript{c}$^{\\mho}$]{gasimova2014}[][\\textsuperscript{a}]{wu2014}[][\\textsuperscript{a}]{coulombe2016b}[][\\textsuperscript{a}]{ye2016}[][\\textsuperscript{a}]{finch2017}[][\\textsuperscript{d}]{orourke2021}[][\\textsuperscript{a}]{newsom2020}[][\\textsuperscript{a}]{coulombe2016}', 
                     '\\parencite[][\\textsuperscript{a}]{timmons2015}', 
                     '\\parencites[][\\textsuperscript{a}]{aydin2014}[][\\textsuperscript{a}]{coulombe2016}',
                     
                     '\\parencites[][\\textsuperscript{b}${\\mho}$]{murphy2011}[][\\textsuperscript{c}$^{\\mho}$]{gasimova2014}[][\\textsuperscript{a}]{wu2014}[][\\textsuperscript{a}]{coulombe2016b}[][\\textsuperscript{a}]{ye2016}[][\\textsuperscript{a}]{finch2017}[][\\textsuperscript{d}]{orourke2021}[][\\textsuperscript{a}]{newsom2020} [][\\textsuperscript{a}]{coulombe2016}[][\\textsuperscript{a}]{aydin2014}', 
                     '', 
                     '\\parencite[][\\textsuperscript{a}]{timmons2015}',  
                     '\\parencite[][\\textsuperscript{a}]{coulombe2016}', 
                     '\\parencites[][\\textsuperscript{b}${\\mho}$]{murphy2011}[][\\textsuperscript{c}$^{\\mho}$]{gasimova2014}[][\\textsuperscript{a}]{wu2014}[][\\textsuperscript{a}]{coulombe2016b}[][\\textsuperscript{a}]{ye2016}[][\\textsuperscript{a}]{finch2017}[][\\textsuperscript{d}]{orourke2021} [][\\textsuperscript{a}]{newsom2020}[][\\textsuperscript{a}]{coulombe2016}', 
                     '\\textbf{Cell 2}', 
                     '\\textbf{Cell 4}', 
                     '\\parencite[][\\textsuperscript{a}]{aydin2014}', 
                     '', 
                     '\\textbf{Cell 6}', 
                     '\\textbf{Cell 8}', 
                     '\\parencite[][\\textsuperscript{a}]{coulombe2016}', 
                     '\\textbf{Cell 11}'),
'Nonlinear pattern' = c('', 
                        '\\parencites[][\\textsuperscript{a}]{timmons2015}[][\\textsuperscript{a}]{finch2017}[][\\textsuperscript{e}$^{\\circ\\triangledown}$]{fine2019}[][\\textsuperscript{e,f}$^{\\triangledown}$]{fine2020}[][\\textsuperscript{g}]{liu2022}[][\\textsuperscript{h}$^{\\mho}$]{liu2021}[][\\textsuperscript{g}$^{\\mho}$]{liu2015}', 
                        
                        '\\parencite[][\\textsuperscript{a}]{timmons2015}', 
                        '\\parencites[][\\textsuperscript{a}$^{\\mho}$]{miller2017}[][\\textsuperscript{g}$^{\\mho}$]{liu2015}', 
                        '\\parencites[][\\textsuperscript{a}]{finch2017}[][\\textsuperscript{e}$^{\\circ\\triangledown}$]{fine2019}[][\\textsuperscript{e,f}$^{\\triangledown}$]{fine2020}[][\\textsuperscript{g}]{liu2022}[][\\textsuperscript{h}$^{\\mho}$]{liu2021}[][\\textsuperscript{g}$^{\\mho}$]{liu2015}[][\\textsuperscript{a}$^{\\mho}$]{miller2017}', 
                        '', 
                        '\\parencite[][\\textsuperscript{a}]{timmons2015}', 
                        
                        '\\textbf{Cell 1 (\\hyperref[Exp3]{Exp. 3})}', 
                        
                        '\\parencites[][\\textsuperscript{a}]{finch2017}[][\\textsuperscript{e}$^{\\circ\\triangledown}$]{fine2019}[][\\textsuperscript{e,f}$^{\\triangledown}$]{fine2020}[][\\textsuperscript{g}]{liu2022}[][\\textsuperscript{h}$^{\\mho}$]{liu2021}',
                        '\\textbf{Cell 3}', 
                        '\\textbf{Cell 5 (\\hyperref[Exp2]{Exp. 2})}' , 
                        
                        '\\parencites[][\\textsuperscript{g}$^{\\mho}$]{liu2015}[][\\textsuperscript{a}$^{\\mho}$]{miller2017}', 
                        '', 
                        '\\textbf{\\centering{\\arraybackslash{Cell 7}}}', 
                        '\\textbf{Cell 9 (\\hyperref[Exp2]{Exp. 2})}', 
                        '\\textbf{Cell 10 (\\hyperref[Exp3]{Exp. 3})}', 
                        '\\textbf{Cell 12}'), check.names = F)
kbl(x = table_1, booktabs = TRUE, format = 'latex', longtable = TRUE, 
  linesep = c('\\cmidrule{1-3}',
        rep(' ', times = 3), '\\cmidrule{1-3}', '\\cmidrule{1-3}', 
        rep(' ', times = 5), '\\cmidrule{1-3}', '\\cmidrule{1-3}',
        rep(' ', times = 3)), 
  align = c('l', 'c', 'c'), 
  caption = 'Summary of Simulation Studies That Have Investigated Longitudinal Issues with Linear and Nonlinear Change Patterns (\\textit{n} = 17)', 
  escape=F) %>%
 column_spec(2, width = '3.5cm') %>%
  column_spec(2, background = c(rep('white', times = 9), 
                                 rep('#E4E2E2', times = 1), 
                                 '#E4E2E2',
                                 'white', 'white',
                                 rep('#E4E2E2', times = 1), 
                                 rep('#E4E2E2', times = 1),
                                 'white',
                                 '#E4E2E2'), 
              width = '8cm') %>%
  column_spec(3, background = c(rep('white', times = 7), 
                                '#C7C4C4',
                                'white', 
                                '#C7C4C4',
                                '#C7C4C4',
                                'white', 'white', 
                                 rep('#C7C4C4', times = 1), 
                                 rep('#C7C4C4', times = 2),
                                 '#C7C4C4'),  
              width = '8cm',  bold = ifelse(grepl(pattern = '^\\d+', x = table_1$Nonlinear),T, F)) %>% 
  
  kable_styling(latex_options= c('hold_position', 'repeat_header'), position = 'left') %>%
  footnote(general = '\\\\textit{Note. }Cells are only numbered for effects that have not been investigated. Cells shaded in light and dark grey indicate effects that have not, respectively, been investigated with linear and nonlinear patterns of change.', 
           general_title = '', footnote_as_chunk = F, symbol_title = '', 
           alphabet_title = '', escape = F, threeparttable = T, 
           alphabet = c('Latent growth curve model. \\\\textsuperscript{b} Second-order latent growth curve model. \\\\textsuperscript{c} Hierarchical Bayesian model. \\\\textsuperscript{d} Bivariate latent change score model. \\\\textsuperscript{e} Functional mixed-effects model. \\\\textsuperscript{f} Nonlinear mixed-effects model. \\\\textsuperscript{g} Bilinear spline model. \\\\textsuperscript{g} Parallel bilinear spline model.'), 
           symbol_manual = c('$\\\\circ$'),
           symbol = c('Manipulated missing data. $^\\\\mho$ Assumed complex error structure (heterogeneous variances and/or correlated residuals). $^\\\\triangledown$ Contained pseudo-time structuredness manipulation.')) %>%
  landscape(margin = '2.54cm') 
```
